### 2016-04-22 writeup/2016-04wikipedia-network/wikipedia-network.acknowledgments.tex
universal template and bibtex format changed

MI is grateful to RJ for pointing out the Reddit post 
\cite{reddit}
describing how the majority of links lead to ``Philosophy,'' inspiring this research.

The authors are also grateful for the suggestions and input provided by Randall Harp.

PSD and CMD acknowledge support from NSF Big Data Grant $\#1447634$.

The authors also acknowledge support from the Vermont Advanced Computing Core which is supported by NASA (NNX 06AC88G), at the University of Vermont for providing High Performance Computing resources that have contributed to the research results reported within this paper.