### 2016-05-02 writeup/2016-04wikipedia-network/correspondence/2016-05-02compsci-biographies.tex
compsci submission

\documentclass[11pt]{article}

\usepackage{textcomp}
\usepackage{graphicx}
\usepackage{url}
\usepackage{hyperref}

\usepackage{marvosym}

\usepackage[svgnames]{xcolor}
%% \definecolor{lightgrey}{rgb}{0.7,0.7,0.7}
%% \definecolor{grey}{rgb}{0.5,0.5,0.5}

\setlength{\oddsidemargin}{20pt}
\setlength{\topmargin}{-50pt}
\setlength{\textwidth}{6in}
\setlength{\textheight}{670pt}

\setlength{\parindent}{0pt}
\setlength{\parskip}{10pt}

\pagestyle{empty}

\renewcommand{\baselinestretch}{1.1}

\newcommand{\etal}{\textit{et al.}}

\begin{document}

\sffamily

\bigskip
\bigskip

Mark Ibrahim is an applied mathematician who's interested in exploring how humans structure and interact with information. Before studying applied mathematics, Mark received a B.A. from Hamilton College where he also studied philosophy, economics, English, and French. Mark has experience in industry assessing market risk using portfolio trading data. He is currently building a project to search videos with speech and a second project serving a network search engine powered by Wikipedia.

Chris Danforth is the Flint Professor of Mathematical, Natural, and Technical Sciences at the University of Vermont. He co-directs the Computational Story Lab, a group of applied mathematicians at the undergraduate, masters, PhD, and postdoctoral level working on large-scale, system problems in many fields including sociology, nonlinear dynamics, networks, ecology, and physics. His research has been covered by the New York Times, Science Magazine, and the BBC among others. Descriptions of his projects are available at his website: http://uvm.edu/∼cdanfort

Peter Sheridan Dodds is a Professor at the University of Vermont (UVM) working on system-level problems in many fields, ranging from sociology to physics. He is Director of UVM's Complex Systems Center, co-Director of UVM's Computational Story Lab, a visiting faculty fellow at the Vermont Advanced Computing Core, and is appointed to the Department of Mathematics and Statistics. He maintains general research and teaching interests in complex systems and networks with a current focus on sociotechnical and psychological phenomena including collective emotional states, contagion, language, and stories. His methods encompass large-scale data collection and analysis, large-scale sociotechnical experiments, and the formulation, analysis, and simulation of theoretical models. Dodds's training is in theoretical physics, mathematics, and electrical engineering with extensive formal postdoctoral and research experience in the social sciences. Dodds's foundational funding was an  NSF CAREER grant awarded by the Social and Economic Sciences Directorate.

%% {\small
%% Professor\\
%% Director of the Vermont Complex Systems Center\\
%% Co-Director, Computational Story Lab\\
%% Vermont Advanced Computing Core\\
%% Department of Mathematics and Statistics\\
%% University of Vermont\\
%%

% $\bullet$ 
% Department of Mathematics and Statistics\\
% $\bullet$ 
% Vermont Advanced Computing Center\\
% $\bullet$ 
% Complex Systems Center\\
% University of Vermont\\
% pdodds@uvm.edu
% $\bullet$
% http://www.uvm.edu/{$\sim$}pdodds

%% \textbf{Suggested reviewers and their relevant expertise:}

\end{document}